Lesson 5
========================================================

### Multivariate Data
We are going to talk about how to examine 3 or more variables at a time

***

### Moira Perceived Audience Size Colored by Age

Tried adding age as color to get a better understanding if age affected whether people are able to guess bettter. But there was too much overplotting.

***

### Third Qualitative Variable
Trying to add a third variable to understand better about the relationships.

```{r Third Qualitative Variable}
library(ggplot2)

pf <- read.csv("../pseudo_facebook.tsv", sep = "\t",  stringsAsFactors=FALSE)

ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)


ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) + 
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)
```

Trying to re-create the same plot by creating the data frame using the `dplyr` package

```{r Third Qualitative Variable 2}
library(dplyr)

pf.fc_by_age_and_gender <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %>%
  ungroup() %>%
  arrange(age)

head(pf.fc_by_age_and_gender)
```

***

### Plotting Conditional Summaries
Making the same plot as earlier using the new data frame which we created just now.

```{r Plotting Conditional Summaries}
ggplot(
  aes(x = age, y = mean_friend_count, color = gender),
  data = pf.fc_by_age_and_gender
) + geom_line()
```

***

### Thinking in Ratios
Having graphs comparing 2 things is great but what if we want to have get a handle on how much is the difference. Let's think in terms of the times of each other.

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Changing format of data so that we have age wise male/female median friend count

```{r}
library(reshape2)
pf.fc_by_age_and_gender.wide <- dcast(pf.fc_by_age_and_gender, 
                                      age ~ gender,
                                      value.var = 'median_friend_count')
```


***

### Ratio Plot
Making a plot to see female users to male users ratio according to age.

```{r Ratio Plot}
ggplot(
    aes(x = age, y = female/male),
    data = pf.fc_by_age_and_gender.wide
) + geom_line() + geom_hline(yintercept = 1, linetype = 2)
```

***

### Third Quantitative Variable
Need to add year joined to the original data frame

```{r Third Quantitative Variable}
pf$year_joined <- floor(2014 - pf$tenure/365)
```

***

### Cut a Variable
Making buckets out of year joined so that each bucket contain comparable number of users

```{r Cut a Variable}
pf$year_joined.bucket <- cut(pf$year_joined, breaks = c(2004, 2009, 2011, 2012, 2014))
```

***

### Plotting it All Together
Adding a plot for the various tenure bucket vs age to see median friend count

```{r Plotting it All Together}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)
```

***

### Plot the Grand Mean
Changing the plot to use mean instead of median. Also want to add overall mean to this as well.

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = mean, linetype = 2)
```

***

### Friending Rate
We want to consider at what rate do people make friends

```{r Friending Rate}
with(subset(pf, tenure >= 1), summary(friend_count/tenure))
```

What is the median friend rate? 0.2205

What is the maximum friend rate? 417.0000 

***

### Friendships Initiated
Do new users go on friending sprees or do old users do this?


```{r Friendships Initiated}
ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, !is.na(year_joined.bucket) & tenure >= 1)) + 
    geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean)
```

***

### Bias-Variance Tradeoff Revisited
Trying to make the graph smoother. Using rounding and smoother

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)


ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_smooth(aes(color = year_joined.bucket))
```

***

### Sean's NFL Fan Sentiment Study

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Making histogram to see how households purchase yogurt.

```{r Histograms Revisited}
yo <- read.csv("yogurt.csv")
str(yo)

yo$id <- factor(yo$id)
str(yo)

qplot(x = price, data = subset(yo, !is.na(price)))
```

***

### Number of Purchases
Getting all number of purchases we introduce a new variable.

```{r Number of Purchases}
yo <- transform(yo, all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)
```

***

### Prices over Time
Notes:

```{r Prices over Time}
qplot(x = all.purchases, data = yo, binwidth = 1)

ggplot(aes(x = time, y = price), data = yo) + geom_jitter(alpha = 1/10, color='orange')
```

***

### Sampling Observations
We sample to understand a subset better before trying to generalize it for everything.

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
library(ggplot2)
set.seed(25121992)

sample.ids <- sample(levels(yo$id),16)
ggplot(aes(time, price),data = subset(yo, yo$id %in% sample.ids)) +
  geom_line()+
  facet_wrap(~id, scales = "free")+
  geom_point(aes(size = all.purchases), pch = 1)
```

***

### The Limits of Cross Sectional Data
Having data at a particular instant of time is great but that does not show trends.

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Let our visualizations help us find which variables to use.

```{r Scatterplot Matrix}
library(GGally)
theme_set(theme_minimal(20))


set.seed(1836)
pf_subset <- pf[, c(2:15)]
ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000), ], axisLabels = 'internal')
```


***

### Even More Variables


***

### Heat Maps
Notes:

```{r Lot more variables}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

