Lesson 3
========================================================

***

### What to Do First?
- What question are you trying to answer?
- Learn about surprises in the data
    - Features that may affect your results
- Understand the variables that are central to your analysis
- Summaries and visualizations of those individual variables

***

### Pseudo-Facebook User Data
- We will be looking at pseudo facebook data
- Similar to the data that Facebook Data Scientists see time to time
- Our goal is to understand User's behaviour and their demographics
    - what they are doing on facebook?
    - what do they use?

```{r Pseudo-Facebook User Data}
getwd()
list.files("..")
pf <- read.csv("../pseudo_facebook.tsv", sep = "\t")
names(pf)
summary(pf)
```

***

### Histogram of Users' Birthdays
We will be using `ggplot` for generating graphs rather than the base graphics package. 

```{r Histogram of Users\' Birthdays}
#install.packages('ggplot2')
 
library(ggplot2)

qplot(x = dob_day, data = pf, bins = 31) + 
  scale_x_continuous(breaks=1:31)
```

`scale_x_continuous` lets us manipulate the x axis

***

#### What are some things that you notice about this histogram?
A lot of people having birthdays on the 1st of the month.

***

### Moira's Investigation
Does people's perception of their audience matches upto the reality? This is important to know because the perception of our audience affects our presentation

How they did it?
- Sent a survey to people pointing them to a post and asked them how many people do you think saw this post?
- counted the actual number of people who had seen the post for at least 1 second
- found that there was a huge difference between people's perceived audience size and their actual audience size

***

### Estimating Your Audience Size
***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
* Moira found that people dramatically understimate their audience size - typically by a factor of 0.25
* Plotted a bunch of histograms of people's guesses

***
### Faceting
Notes:

```{r Faceting}

```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}

```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}

```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}

```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}

```

#### Who on average has more friends: men or women?
Response:

#### What's the difference between the median friend count for women and men?
Response:

#### Why would the median be a better measure than the mean?
Response:

***

### Tenure
Notes:

```{r Tenure}

```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}

```

***

### Labeling Plots
Notes:

```{r Labeling Plots}

```

***

### User Ages
Notes:

```{r User Ages}

```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:

***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}

```

***


### Frequency Polygons

```{r Frequency Polygons}

```

***

### Likes on the Web
Notes:

```{r Likes on the Web}

```


***

### Box Plots
Notes:

```{r Box Plots}

```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}

```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}

```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}

```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}

```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!