Lesson 2 - R Basics
========================================================

### 01 - The Power of R
* Advantage of using R to perform Data Analysis
* Basic data types and commands of R

### Resources
* [Foodbourne Chicago finds Dodgy Restaurants with Tweets](http://blog.revolutionanalytics.com/2013/08/foodborne-chicago.html)
* [Sensitivity and Specificity and ROC Curves for Binary Classifiers](http://blog.yhathq.com/posts/roc-curves.html) (more advanced reading; focus on understanding TPR and FPR). 

### 02 - The Power of R
* Used `textcat` package for analyzing tweets

### 03 - Why R
* We can build our analysis line by line
* Having R scripts allows you to easily share your work with others
* Has over 2000 user contributed packages
* Has large, active and growing community of users

Let's see any example of what we can get out of R using a few lines of code

```{r Example Scatterplot}
#install.packages('ggplot2', dependencies = TRUE) 
library(ggplot2) 

#install.packages('RColorBrewer', dependencies = TRUE) 
library(RColorBrewer) 

data("diamonds")

qplot(data = diamonds, x = carat, y = price, color = cut) +
  scale_color_brewer(palette = 'Accent')
```

### 08 - Demystifying R

Go through `demystifying.R` to try out R.

### 10 - Getting Help
* Using `?FUNCTION` to get help for any function

#### Resources
* [Quick R]('http://www.statmethods.net/')
* [R Cookbook]('http://www.cookbook-r.com/')
* [R-Bloggers]('http://www.r-bloggers.com/')
* [StackOverflow About R]('http://stackoverflow.com/tags/r/info')
* [StackOverflow R FAQ]('http://stackoverflow.com/questions/tagged/r-faq%20')
* [Google's R Style Guide]('https://google.github.io/styleguide/Rguide.xml')

### 11 - Read and Subset Data

```{r Check and set current working directory}
getwd()
#setwd('')
```

```{r Reading in data}
statesInfo <- read.csv('./stateData.csv')
```
* Double clicking the data frame in Environment tab allows you to view the data in a tabular format in RStudio
* We can subset a dataframe to select some rows based on some conditions passed to `subset` function
* We can see the number of rows and columns using `dim`
```{r Subsetting data}
stateSubset <- subset(statesInfo, state.region == 1)
head(stateSubset, 2)
dim(stateSubset)

stateSubsetBracket <- statesInfo[statesInfo$state.region == 1, ]
head(stateSubsetBracket, 2)
dim(stateSubsetBracket)
```

#### Resources
* Read this [fantastic tutorial](http://flowingdata.com/2015/02/18/loading-data-and-basic-formatting-in-r/) on loading data into R by Nathan Yau.
* [How to Subset Data](http://www.statmethods.net/management/subset.html)

### 12 - R Markdown Documents
* Go through `What_is_a_RMD_file.Rmd` file to see how a RMD file looks like
* The second file, `demystifyingR2.Rmd`, is the file that Chris mentions at the end of the video. Work through this file and use it to answer the next quiz! 

You'll need to install and load the package knitr in order to use the KNIT HTML button . Run the following commands in the RStudio console to install and load knitr. 

```{r Install Knitr to make HTML}
#install.packages('knitr', dependencies = T) 
library(knitr)
```

### 13 - R Markdown Documents
Which cars have `mpg` greater than or equal to 30 or `hp` less than 60?

```{r}
data("mtcars")
str(mtcars)
subset(mtcars, mpg >= 30 | hp < 60)
```

### 14 - Factor Variables

We will look at some data collected from survey of reddit users. This survey asked users about
    + Demographics - gender, nationality, employment
    + what type of cheese they would be
    + prefer dogs, cats or turtles

```{r Reading reddit data and seeing some details}
reddit <- read.csv("reddit.csv")
str(reddit)
```

How many people are in each group of employment status
```{r}
table(reddit$employment.status)
```

We can also get these counts and other data points using the `summary` function
```{r}
summary(reddit)
```

#### Resources
* Learn more about [data types](http://www.statmethods.net/input/datatypes.html).
* [Who is Reddit?](http://www.redditblog.com/2011/09/who-in-world-is-reddit-results-are-in.html)

### 15 - Ordered Factors

Looking more closely at the facotr variables. We can look at the level of factor variables using the following

```{r}
levels(reddit$age.range)
```

Let's create a plot to see how many users are in each range

```{r}
library(ggplot2)
qplot(data = reddit, x = age.range)
```

Now the graph is in order but the 'Under 18' group should have been the first one. For this to happen we want ordered factors.

### 18 - Setting Levels of Ordered Factors

The problem is much worse in some other cases. 

```{r}
qplot(data = reddit, x = income.range)
```

Setting the order can be done like this
```{r}
reddit$age.range <- ordered(reddit$age.range, 
                            levels = c('Under 18', '18-24', '25-34', '35-44', '45-54',
                                       '55-64', '65 or Above'))

qplot(data = reddit, x = age.range)

#Alternate way to set order of factors
reddit$age.range <- factor(reddit$age.range, 
                            levels = c('Under 18', '18-24', '25-34', '35-44', '45-54',
                                       '55-64', '65 or Above'),
                           ordered = T)

qplot(data = reddit, x = age.range)
```


#### Resources
* [Learn how to set and order factor levels.](http://statistics.ats.ucla.edu/stat/r/modules/factor_variables.htm)


### 19 - Data Munging
* Data sets used so far were tidy
* Normally we won't have tidy data
* Or we may have to reshape or re arrange data into different formats

#### Resources
* [Data Wrangling with MongoDB: Data Manipulation and Retrieval](https://www.udacity.com/course/ud032)
* Read this [fantastic tutorial](http://flowingdata.com/2015/02/18/loading-data-and-basic-formatting-in-r/) on loading data into R by Nathan Yau.
* [Tidy Data](http://vita.had.co.nz/papers/tidy-data.pdf) by Hadley Wickham
* [Tidy Data Presentation](http://courses.had.co.nz.s3-website-us-east-1.amazonaws.com/12-rice-bdsi/slides/07-tidy-data.pdf) by Hadley Wickham (June 2012)

### 20 - Advice for Data Scientists
* Find data sets that you are interested in
* Work with them, play with the data
* Good data science come from good questions, not fancy techniques
* Motivating the research with an idea that you care about

#### Resources
* [Eytan Bakshy](http://eytan.github.io/)
* [Sean Taylor](http://seanjtaylor.com/)